---
title: "Data 607 Project 3"
output: html_document
author: "Inna Yedzinovich, Zaneta Paulusova, Md Asaduzzaman, Md.Asadul Islam,Md.Simon Chowdhury,"
date: "`r Sys.Date()`"

---
## Introduction 

TDODOOOOOOOOOOOOOO


#### Tables creation for our data

```{r}
library(DBI)
library(RMySQL)

mysqlconnection = dbConnect(RMySQL::MySQL(),
                            dbname='project3',
                            host='localhost',
                            port=3306,
                            user='root',
                            password='@Helih254')

dbListTables(mysqlconnection)

experience_result = dbSendQuery(mysqlconnection, "select * from experience")
experience.data.frame = fetch(experience_result)
print(experience.data.frame) 

interestareas_result = dbSendQuery(mysqlconnection, "select * from interestareas")
interestareas.data.frame = fetch(interestareas_result)
print(interestareas.data.frame) 

learningresources_result = dbSendQuery(mysqlconnection, "select * from learningresources")
learningresources.data.frame = fetch(learningresources_result)
print(learningresources.data.frame)

programminglanguages_result = dbSendQuery(mysqlconnection, "select * from programminglanguages")
programminglanguages.data.frame = fetch(programminglanguages_result)
print(programminglanguages.data.frame)

respondents_result = dbSendQuery(mysqlconnection, "select * from respondents")
respondents.data.frame = fetch(respondents_result)
print(respondents.data.frame)

softskills_result = dbSendQuery(mysqlconnection, "select * from softskills")
softskills.data.frame = fetch(softskills_result)
print(softskills.data.frame)

valuableskills_result = dbSendQuery(mysqlconnection, "select * from valuableskills")
valuableskills.data.frame = fetch(valuableskills_result)
print(valuableskills.data.frame)

```

WRITE BETTER TODOOOO

Load the CSV into R
Tidy and Normalize the Data
Load the Data into the Database

```{r}
library(dplyr)
library(tidyr)

data <- read.csv("https://raw.githubusercontent.com/simonchy/DATA607/refs/heads/main/week%208/Cleaned_Augmented_Data_Science_Skills.csv")
colnames(data)
colnames(data) <- tolower(colnames(data))

data <- data %>%
  rename(
    timestamp = timestamp,
    first_name = first.name.or.nickname,
    valuable_skills = list.the.5.most.valuable.data.science.skills..separated.by.commas.,
    email = email.address,
    age = age,
    data_science_experience = any.data.science.data.analytics.experience.,
    software_engineering_experience = any.software.engineering.experience.,
    programming_languages = which.programming.languages.do.you.use.most.frequently.,
    learning_resources = what.resources.do.you.use.for.learning.new.data.science.skills.,
    interest_areas = what.areas.of.data.science.are.you.most.interested.in.learning.more.about.,
    skill_1 = name..1.most.most.valuable.data.science.skill,
    skill_2 = name..2.most.most.valuable.data.science.skill,
    skill_3 = name..3.most.most.valuable.data.science.skill,
    skill_4 = name..4.most.most.valuable.data.science.skill,
    skill_5 = name..5.most.most.valuable.data.science.skill,
    soft_skill = which.soft.skill.do.you.think.is.most.important.for.a.data.scientist.
  ) %>%
  mutate(across(everything(), tolower))

# Create respondent_id before separating rows
data <- data %>%
  mutate(respondent_id = row_number())

# Normalize the data
respondents <- data %>%
  select(first_name, age, respondent_id)

experience <- data %>%
  select(data_science_experience, software_engineering_experience, respondent_id) %>%
  mutate(experience_id = row_number())

valuable_skills <- data %>%
  select(skill_1, skill_2, skill_3, skill_4, skill_5, respondent_id) %>%
  pivot_longer(cols = starts_with("skill_"), names_to = "skill_rank", values_to = "skill_name") %>%
  mutate(valuable_skill_id = row_number())

programming_languages <- data %>%
  select(respondent_id, programming_languages) %>%
  separate_rows(programming_languages, sep = ",") %>%
  mutate(language_id = row_number())

learning_resources <- data %>%
  select(respondent_id, learning_resources) %>%
  separate_rows(learning_resources, sep = ",") %>%
  mutate(resource_id = row_number())

interest_areas <- data %>%
  select(respondent_id, interest_areas) %>%
  separate_rows(interest_areas, sep = ",") %>%
  mutate(interest_id = row_number())

soft_skills <- data %>%
  select(soft_skill, respondent_id) %>%
  mutate(soft_skill_id = row_number())

#conn <- dbConnect(RMySQL::MySQL(), user = Sys.getenv("MYSQL_USER"), password = #Sys.getenv("MYSQL_PASSWORD"), host = Sys.getenv("MYSQL_HOST"), dbname = #Sys.getenv("MYSQL_DBNAME"))

# Load the data into the database
#dbWriteTable(conn, "respondents", respondents, overwrite = TRUE, row.names = FALSE)
#dbWriteTable(conn, "experience", experience, overwrite = TRUE, row.names = FALSE)
#dbWriteTable(conn, "valuableskills", valuable_skills, overwrite = TRUE, row.names = FALSE)
#dbWriteTable(conn, "programminglanguages", programming_languages, overwrite = TRUE, row.names = FALSE)
#dbWriteTable(conn, "learningresources", learning_resources, overwrite = TRUE, row.names = FALSE)
#dbWriteTable(conn, "interestareas", interest_areas, overwrite = TRUE, row.names = FALSE)
#dbWriteTable(conn, "softskills", soft_skills, overwrite = TRUE, row.names = FALSE)

#dbDisconnect(conn)

```

BETTER DESCRIPTION TODOOOOO
Load the data from the database.
Count the occurrences of each skill.
Create a bar plot using ggplot2.

```{r}
library(ggplot2)

conn <- dbConnect(RMySQL::MySQL(), user = Sys.getenv("MYSQL_USER"), password = Sys.getenv("MYSQL_PASSWORD"), host = Sys.getenv("MYSQL_HOST"), dbname = Sys.getenv("MYSQL_DBNAME"))
valuable_skills <- dbReadTable(conn, "valuableskills")
dbDisconnect(conn)

skill_counts <- valuable_skills %>%
  count(skill_name, sort = TRUE)
print(skill_counts)

ggplot(skill_counts, aes(x = reorder(skill_name, n), y = n)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Valuable Data Science Skills", x = "Data Science Skills", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Let's make it look cleaner: 

```{r}

# Group similar skills together using regex grepl
valuable_skills <- valuable_skills %>%
  mutate(skill_group = case_when(
    grepl("machine learning|ML", skill_name, ignore.case = TRUE) ~ "Machine Learning",
    grepl("programming|coding|software|Python|R", skill_name, ignore.case = TRUE) ~ "Programming",
    grepl("statistics|statistical", skill_name, ignore.case = TRUE) ~ "Statistics",
    grepl("data|database", skill_name, ignore.case = TRUE) ~ "Data Management",
    grepl("math|algorithm", skill_name, ignore.case = TRUE) ~ "Mathematics & Algorithms",
    grepl("communication|presentation", skill_name, ignore.case = TRUE) ~ "Communication",
    grepl("teamwork|collaborat", skill_name, ignore.case = TRUE) ~ "Teamwork",
    grepl("critical thinking|problem solving|analysis", skill_name, ignore.case = TRUE) ~ "Critical Thinking",
    TRUE ~ "Other"
  ))

skill_counts <- valuable_skills %>%
  count(skill_group, sort = TRUE)

ggplot(skill_counts, aes(x = reorder(skill_group, n), y = n)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Valuable Data Science Skills", x = "Skill Categories", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


MORE ANALYSIS

TODOOOOOOOOO


### Conclusion

TODOOOOOOOOOOOO